Supplementary Information

### 2020-10-16 doc/manuscript/MEE_manuscript.Rmd
bit of manuscript fix-up

### Length planning:
## Main text ~4300 words --> ~28 paragraphs averaging 150 words
# Intro- 900/ 6
# Methods- 1500 / (11)
# Results - 1000/ (6)
# Discussion - 900 / (5)
## References: figure ~40 words each, so 50 refs = 2000
## Display items: figure ~100 words for each caption = 700 words
# Table1- table of sites
# Fig1- schematic illustrating our process
# Fig2- climate senstivity (SCBI example): (a) Helcoski quilt plot, (b) Climwin panel a, (c) GLS output
# Fig3-  multipanel- GLS results for climate sensitivity at all sites (3 columns with the 3 variable type groupings, 1 row per site) 
# Fig4-relationship of ∆r and ∆AGB to DBH (all site-species combinations)
# Fig5/ table2- DBH/climate interactions

```

@@

```{r eval=FALSE, echo=FALSE}

# I tried to put in table, as in [Camille's paper](https://github.com/EcoClimLab/ForestGEO-tree-size-carbon/blob/master/paper_treesize_draft.Rmd), but couldn't make it work.
library(knitr)
library(kableExtra)
sites <- read.csv("tables_figures/sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
kable(sites)
      #caption = "Study sites")
```

**Table 1 | Sites included in this analysis**. Here and throughout, sites are ordered by latitude. Additional site information is provided in Table S1, and species and their sample sizes are detailed in Tables S2-S3. *Climate will be filled in based on CRU data.*
![ ](tables_figures/sites_table.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @vlam_temperature_2014; @maxwell_declining_2016; @sniderhan_growth_2016; @alfaro-sanchez_growth_2017; @helcoski_growing_2019; @aus_de_ar_tree_2018; @bumann_assessing_2019; @samonil_individual-based_2013; @kaspar_species-specific_nodate
---

@@

Our analysis consisted of two main steps: (1) identification of the most important climate drivers, and (2) combining $DBH$ and climate drivers into a multivariate model (Fig. **1**). The analysis was run separately for each site and each response variable ($\Delta r$, $BAI$, or $\Delta AGB$).

@@

**We show that DBH is a strong driver of growth (any metric) relative to climate and that climate sensitivity often varies with DBH-- findings that have important implications for understanding changes in forest productivity in response to climate change.**
The observed importance of $DBH$ as a driver of tree growth reinforces the concept that changes in forest structure/ demography are critical to changes in forest productivity [REFS; @mcdowell_pervasive_2020].
The prevalence of interactive $DBH$-climate effects reinforces the concepts that large trees tend to be disproportionately sensitive to drought  [@bennett_larger_2015; @mcgregor_tree_nodate] but less sensitive to high temperatures *(Rollinson et al. in review)*, implying that size should be considered in tree-ring analyses of climate sensitivity.
However, the relatively strong importance of $DBH$ in no way negates the importance of climate, as its effect would dominate changes in productivity in stands with relatively constant size structure.
Some studies have made a start at combining tree rings and forest census data to get at the climate sensitivity of whole-forest productivity [e.g., @helcoski_growing_2019], but have been limited in that they were not underlain by models of individual tree growth that simultaneously characterized the effects of tree size and climate, as we do here.
Further work will be required to combine this model with forest census data in order to understand the climate sensitivity of whole-forest productivity.

>>>>

### Length planning:
## Main text ~4300 words --> ~28 paragraphs averaging 150 words
# Intro- 900/ 6
# Methods- 1500 / (11)
# Results - 1000/ (6)
# Discussion - 900 / (5)
## References: figure ~40 words each, so 50 refs = 2000
## Display items: figure ~100 words for each caption = 700 words
```

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
sites = 1
#Figures
schematic=1
SCBI_example=2
clim_sens=3
size=4
# Fig5/ table2- DBH/climate interactions
```

@@

```{r sites, eval=FALSE, echo=FALSE}
# This generates a table, but doesn't format nicely and messes up special characters in names, so turned off.
library(knitr)  
#kableExtra  library will mess it up!
sites <- read.csv("tables_figures/sites.csv", stringsAsFactors = FALSE, check.names = FALSE)

knitr::kable(sites[, c("site code",	"site name",	"location",	"July T",	"Jan T",	"Ann. Precip.",	"vegetation type(s)",	"n species",	"n cores",	"original publication(s)")], 
                  col.names = c("site code",	"site name",	"location",	"July T",	"Jan T",	"Ann. Precip.",	"vegetation type(s)",	"n species",	"n cores",	"original publication(s)"), 
                  caption = "Table 1. Sites included in this analysis.") 
```

**Table 1 | Sites included in this analysis**. Here and throughout, sites are ordered by mean July T. Additional site information is provided in Table S1, and species and their sample sizes are detailed in Tables S2-S3. *Climate will be filled in based on CRU data.*
![ ](tables_figures/sites_table.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @vlam_temperature_2014; @maxwell_declining_2016; @sniderhan_growth_2016; @alfaro-sanchez_growth_2017; @helcoski_growing_2019; @aus_de_ar_tree_2018; @bumann_assessing_2019; @samonil_individual-based_2013; @kaspar_species-specific_nodate
---

@@

Our analysis consisted of two main steps: (1) identification of the most important climate drivers, and (2) combining $DBH$ and climate drivers into a multivariate model (Fig. `r schematic`). The analysis was run separately for each site and each response variable ($\Delta r$, $BAI$, or $\Delta AGB$).

@@

**We show that DBH is a strong driver of growth (any metric) relative to climate and that climate sensitivity often varies with DBH-- findings that have important implications for understanding changes in forest productivity in response to climate change.**
The observed importance of $DBH$ as a driver of tree growth reinforces the concept that changes in forest structure/ demography are critical to changes in forest productivity [REFS; @mcdowell_pervasive_2020].
The prevalence of interactive $DBH$-climate effects reinforces the concepts that large trees tend to be disproportionately sensitive to drought  [@bennett_larger_2015; @mcgregor_tree_nodate] but less sensitive to high temperatures *(Rollinson et al. in review)*, implying that size should be considered in tree-ring analyses of climate sensitivity.
(@hacket-pain_consistent_2016 show an interaction)
However, the relatively strong importance of $DBH$ in no way negates the importance of climate, as its effect would dominate changes in productivity in stands with relatively constant size structure.
Some studies have made a start at combining tree rings and forest census data to get at the climate sensitivity of whole-forest productivity [e.g., @helcoski_growing_2019], but have been limited in that they were not underlain by models of individual tree growth that simultaneously characterized the effects of tree size and climate, as we do here.
Further work will be required to combine this model with forest census data in order to understand the climate sensitivity of whole-forest productivity.
